### 2021-12-22 paper/arXiv/paper.tex
add .gitignore and add paper vital files
.gitignore will remove certain folder which contains too many files. With this, I found the vital paper files and updated them. Also reorganized files to make three versions of the paper: for elsevier (which I've been working with) and for MDPI (deadline at jan-31 2022) and arXiV.org (without format).

\documentclass[12pt]{article}

\title{Art\'iculo}
\author{Jorge Tarango-Yong}

\begin{document}
\maketitle

Cuerpo del texto aqu\'i
\end{document}