### 2021-12-22 paper/meteor-properties.tex

\subsection{Meteoroid masses}

Raul should write this section